---
title: "Showing off some ggplot2 skills"
author: "Team 09: Cyrus Tadjiki and Matt McCoy."
date: </br>`r format(Sys.time(), '%d %B %Y')`
output: github_document
---

```{r setup, include=FALSE}
## These next lines set the default behaviour for all R chunks in the .Rmd document.
## I recomend you take a look here: https://rmarkdown.rstudio.com/authoring_rcodechunks.html
knitr::opts_chunk$set(echo = TRUE) ## Show all R output
knitr::opts_chunk$set(cache=TRUE) ## Cache the results to increase performance.
```

## Goal and requirements

The goal of this assignment is simple: I want you to produce *two* different figures using **ggplot2**. They don't have to be identify causal relationships, or anything like that. I just want you to stretch your visualization legs and demonstrate any new (or existing) **ggplot2** skills that you have acquired since our first lecture. **Important:** Both team members should contribute equally to this assignment (e.g. repsosible for one figure each).

Some additional points:

- You are free to use any dataset that comes in-built with base R, or bundled together with an external R package. See [here](https://vincentarelbundock.github.io/Rdatasets/datasets.html) for an impressive list.

- That being said, I would especially encourage you to use your own data. 
  - I know we haven't gotten to data importation yet, but take a look [here](https://support.rstudio.com/hc/en-us/articles/218611977-Importing-Data-with-RStudio) if you need help. I would recommend that you install the **readr**, **read_excel** and **haven** packages first, though.
  - If your dataset isn't proprietary (or isn't being read directly off the web), please save it in the (empty) `data/` folder of this repo. 

- You can use the same dataset for all four of your plots. Or you can use a new dataset for each plot. Regardless of what you choose, I want you to try and use different geoms for each figure.

- Any other **ggplot2** skills and add-ons like faceting, changing aesthetic scales or legends, using different themes (e.g. from the **ggthemes** package), animation, etc. are all welcome and encouraged.

- I want to *see* the code that produces the figures. (Don't use `echo=FALSE` in any of the code chunks, if that means anything to you.)

### What you will be graded on

- Are your figures clear? (e.g. lack of chart chunk, non-overlapping labels)
- Are your figures compelling? (e.g. use an appropriate geom for the insight that you want to convey)
- Variation (I don't want to see four line charts of the same dataset... Be creative)
- Did you read and follow my instructions (e.g. describe your data and figures, show the code that produced the figures, include data in the `/data` folder, etc)
- etc.

Lastly, don't forget to knit the assignment (click the "Knit" button, or press `Ctrl+Shift+K`) before submitting!

## Start the assignment

Here is a R code chunk for you to load your packages. Feel free to load as many other packages and insert as many additional code chunks (`Ctr+Alt+I`) as you need. You can always load additional packages in the their own chunk blocks, but I recommend loading all of your packages together at the top.)

```{r libs}
if (!require("pacman")) install.packages("pacman")
#Install Ecdata if not already installed as it contains the dataset.
if (!require("Ecdat")) install.packages("Ecdat")
## Note: The `p_load()` function from the pacman package is a convenient way to 
## install (if necessary) and load packages all at once. You can think of this
## as an alternative to the normal `library(ggplot2); library(here);...` way of
## loading R packages.
pacman::p_load(ggplot2, here, Ecdat, dplyr)
```

```{r data1}
#load data
data(Mathlevel)
```

```{r stats1}
# quick summary statistics

summary(Mathlevel)

head(Mathlevel, n=10)

tail(Mathlevel, n=5)

```


### Figure 1: Title description

Briefly describe the data here. Then produce the figure.
**Here is my description**

We found the data from the Vincent Arel-Bundock github, and it is located in the Ecdat package. This study collected data on the level of calculus that had been attained by students taking Advanced Microeconomics.  

```{r fig1}
# Plot code here
```

Finally, tell me what I am looking at. Point out anything that you think is of interest. Now do the same for Figs. 2 below...

### Figure 2: Title description

```{r fig2}
# Plot code here

ggplot(Mathlevel) +
  aes(x = sat, y = major, fill = major) +
  geom_boxplot(outlier.color="red") 

```
