---
title: "Showing off some ggplot2 skills"
author: "Team 09: Cyrus Tadjiki and Matt McCoy."
date: </br>`r format(Sys.time(), '%d %B %Y')`
output: github_document
---

```{r setup, include=FALSE}
## These next lines set the default behaviour for all R chunks in the .Rmd document.
## I recomend you take a look here: https://rmarkdown.rstudio.com/authoring_rcodechunks.html
knitr::opts_chunk$set(echo = TRUE) ## Show all R output
knitr::opts_chunk$set(cache=TRUE) ## Cache the results to increase performance.
```

## Goal and requirements

The goal of this assignment is simple: I want you to produce *two* different figures using **ggplot2**. They don't have to be identify causal relationships, or anything like that. I just want you to stretch your visualization legs and demonstrate any new (or existing) **ggplot2** skills that you have acquired since our first lecture. **Important:** Both team members should contribute equally to this assignment (e.g. repsosible for one figure each).

Some additional points:

- You are free to use any dataset that comes in-built with base R, or bundled together with an external R package. See [here](https://vincentarelbundock.github.io/Rdatasets/datasets.html) for an impressive list.

- That being said, I would especially encourage you to use your own data. 
  - I know we haven't gotten to data importation yet, but take a look [here](https://support.rstudio.com/hc/en-us/articles/218611977-Importing-Data-with-RStudio) if you need help. I would recommend that you install the **readr**, **read_excel** and **haven** packages first, though.
  - If your dataset isn't proprietary (or isn't being read directly off the web), please save it in the (empty) `data/` folder of this repo. 

- You can use the same dataset for all four of your plots. Or you can use a new dataset for each plot. Regardless of what you choose, I want you to try and use different geoms for each figure.

- Any other **ggplot2** skills and add-ons like faceting, changing aesthetic scales or legends, using different themes (e.g. from the **ggthemes** package), animation, etc. are all welcome and encouraged.

- I want to *see* the code that produces the figures. (Don't use `echo=FALSE` in any of the code chunks, if that means anything to you.)

### What you will be graded on

- Are your figures clear? (e.g. lack of chart chunk, non-overlapping labels)
- Are your figures compelling? (e.g. use an appropriate geom for the insight that you want to convey)
- Variation (I don't want to see four line charts of the same dataset... Be creative)
- Did you read and follow my instructions (e.g. describe your data and figures, show the code that produced the figures, include data in the `/data` folder, etc)
- etc.

Lastly, don't forget to knit the assignment (click the "Knit" button, or press `Ctrl+Shift+K`) before submitting!

## Start the assignment

Here is a R code chunk for you to load your packages. Feel free to load as many other packages and insert as many additional code chunks (`Ctr+Alt+I`) as you need. You can always load additional packages in the their own chunk blocks, but I recommend loading all of your packages together at the top.)

```{r libs}
if (!require("pacman")) install.packages("pacman")
#Install Ecdata if not already installed as it contains the dataset.
if (!require("Ecdat")) install.packages("Ecdat")
#Install gridExtra if not already installed as it contains the dataset.
if (!require("gridExtra")) install.packages("gridExtra")
## Note: The `p_load()` function from the pacman package is a convenient way to 
## install (if necessary) and load packages all at once. You can think of this
## as an alternative to the normal `library(ggplot2); library(here);...` way of
## loading R packages.
pacman::p_load(ggplot2, here, Ecdat, dplyr, gridExtra)
```

```{r data1}
#load data
data(Mathlevel)
```

```{r stats1}
# quick summary statistics

summary(Mathlevel)

head(Mathlevel, n=10)

tail(Mathlevel, n=5)

```
### **Description of Data**

We found the data from the Vincent Arel-Bundock github, and it is located in the Ecdat package. This study collected data on the level of calculus that had been attained by students taking Advanced Microeconomics from 1983 to 1986. The students are also categorized by their major. It appears to have been a fairly small cross-sectional study as there are only 609 observations. 

**Brief note on this data set:** 
It was very frustrating making plots of data when the data set "MathLevel" and a important variable "mathlevel" had the same exact spelling and only on variation of the first letter being capitalized or not. R wasn't always sure what todo so I had to specify with the pipeline %>% occasionally. In the future I would've avoided this by making my own .csv and renaming the columns or mutating the data within R. 

###  Fiqure 1: Distribution of forgien langue proficiency and sex with Math performance
```{r fig1}

Gender_plot <- ggplot(Mathlevel) +
                aes(y = mathlevel, x = sat, color = sex) +
                geom_jitter() + labs(title="Math Performance and Sex", x = "SAT Math Score", y = "Highest Math Class Taken") +
                scale_color_discrete(name = "Sex", labels = c("Male", "Female")) + 
                theme_dark() +
                theme(legend.position="bottom")

Language_plot <- ggplot(Mathlevel) +
                  aes(y = mathlevel, x = sat, color = language) +
                  geom_jitter() + labs(title="Math Performance and Language", x = "SAT Math Score", y = "Highest Math Class Taken") +
                  scale_color_discrete(name = "Foreign Language", labels = c("No", "Yes")) + 
                  theme_dark() +
                  theme(legend.position="bottom")
```

```{r}
grid.arrange(Gender_plot, Language_plot, ncol = 2)
```

## Description of **Figure 1**:
I plotted two side-by-side geom_jitter() plots to look and see if there is a trend between SAT Math Scores and highest level of math attained in College. I further more duplicated the plots and colored sex in the first one and foreign language proficiency in the other to see if these binary variables are clustered towards the top or bottom of these jitters. My parents really pushed me to take Spanish in middle school and told me it would help me be better rounded at math because the use the same hemisphere of the brain. I though that was nonsense and my plots help support that it was nonsense showing no cluster in language. Additionally, there are no clusters with Sex and math performance. I believe the reason we get three distributed chunks of observations is because of the different requirements of math for specific majors. Non-STEM majors most likely belong in the lowest chunk and Economics and Natural Science Majors most likely belong in the highest chunk due to those majors requiring the highest math requirements. 


### Figure 2: Distribution of SAT Math scores By Major

```{r fig2}
# Plot code here

ggplot(Mathlevel) +
  aes(x = sat, y = major, fill = major) +
  geom_boxplot(outlier.color="red") + labs(title = "SAT Math Score Distribution By Major", x = "SAT Math Score", y = "Major", fill = "Major") 

```
This Boxplot depicts the distribution of SAT math scores within each of the majors. I was curious to see if there would be any correlation between specific majors and higher math scores. It does not appear that math scores influence what major someone picks as each major has a very similar distribution and interquartile range. The Natural Sciences does have an edge on the upper whisker indicating that their top 25% of scores are slightly higher. The Economics and Other majors had a slightly worse lower whisker score indicating there are slightly more low scores. The outliers are highlighted as a red dot. It's interesting that there are no outliers in the Humanities major, and they have the most compact distribution of scores. The spread of math scores is fairly symmetrical between group, and the data is not skewed in either direction.     